---
title: "censusr"
author: ""
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{censusr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r setup, echo = FALSE}
knitr::opts_chunk$set(eval = F)
```

The Census has made a very nice API for data scientists to access their data
tables. The `censusr` package will help R users access this API in a convenient
and R-like way.

The API works by sending a specially-formatted URL to the the Census API server,
which returns an XML or JSON document containing the requested information. In 
practice, any table available on American FactFinder is available through the
API, though the user will need to find the raw name for the variable in the
Census API 
[guide](http://www.census.gov/data/developers/data-sets/acs-survey-5-year-data.html).

# Setup
These instructions are modified from hadley's 
[API best practices documentation](https://cran.r-project.org/web/packages/httr/vignettes/api-packages.html).

1. Users of this package will need to request an API key, which is available for 
free from the Census Bureau on request. Go to
[http://api.census.gov/data/key_signup.html](http://api.census.gov/data/key_signup.html)
to register. Copy this token to your clipboard.

2. Identify your home directory. If you are not sure what it is, enter 
`normalizePath("~/")` in an R session. If in RStudio, use the R console.

3. Create a new text file. If in RStudio, do File > New File > Text file.

4. Create a line like this:

```
CENSUS_TOKEN=blahblahblahblahblahblah

```

where the name `CENSUS_TOKEN` reminds you which API this is for and 
`blahblahblahblahblahblah` is your token, pasted from the clipboard. Make sure 
the last line in the file is empty. (If it is not empty, R will silently fail 
to load the file. If you're using an editor that shows line numbers, there 
should be two lines, where the second one is empty.)

5. Save this file in your home directory with the filename `.Renviron`. If 
questioned, YES you do want to use a filename that begins with a dot `.`.

Note that by default dotfiles are usually hidden. But within RStudio, the file 
browser will make `.Renviron` visible and therefore easy to edit in the future.

6. Restart R. `.Renviron` is processed only at the start of an R session.

7. Use `Sys.getenv()` to access your token. For example,

```{r library, eval=F}
call_census_api(..., api_key = Sys.getenv("CENSUS_TOKEN") ...)
```

FAQ: Why define this environment variable via `.Renviron` instead of in 
`.bash_profile` or `.bashrc`?

Because there are many combinations of OS and ways of running R where the 
`.Renviron` approach "just works"" and the bash stuff does not. When R is a 
child process of, say, Emacs or RStudio, you can't always count on environment 
variables being passed to R. Put them in an R-specific start-up file and save 
yourself some grief.

# Use
The package works by sending a list of requested variables and a list of
geographies. The call below requests the number of households owning 0, 1, 2, 3,
or 4 or more vehicles in Wake County, North Carolina (`geoid = 37183`). We
specify that we want this table for 2012 5-year summary level.

```{r censustables, purl = FALSE}
library(censusr)
call_census_api(
  paste("B08201_", sprintf("%03d", 2:6),  "E", sep = ""),
  names = c(0:4), geoids = "37183",  
  data_source = "acs", year =  2012, period = 5) 
```

We can use the `allgeos` argument to say that we actually want these variables
for *all* census tracts within Wake County.

```{r allgeos, purl = FALSE}
est <- call_census_api(
  paste("B08201_", sprintf("%03d", 2:6),  "E", sep = ""),
  names = paste0("est_", c(0:4)), geoids = "37183",  allgeos = "tr",
  data_source = "acs", year =  2012, period = 5) 
```

If we want the margins of error on this table instead of the estimates, we can
change the variable to call the `M` type instead of the `E` type.

```{r margins, purl=FALSE}
moe <- call_census_api(
  paste("B08201_", sprintf("%03d", 2:6),  "M", sep = ""),
  names = paste0("moe_", c(0:4)), geoids = "37183",  allgeos = "tr",
  data_source = "acs", year =  2012, period = 5) 
```
